Independent Component Analysis

### 2022-04-30 paper.tex
add ex2

% You should use BibTeX and apsrev.bst for references
% Choosing a journal automatically selects the correct APS
% BibTeX style file (bst file), so only uncomment the line
% below if necessary.
%\bibliographystyle{apsrev4-2}
\usepackage{graphicx}
\usepackage{amsmath,amsthm,amssymb,mathtools}
\usepackage{amsfonts}
\usepackage{bm}
\usepackage{color}
\frenchspacing
%\graphicspath{{~/data/KannanLu/exp623/article/}}

@@

\subsection{FastICA}
\subsection{alternative objective function, skewness and kurtosis (Kannan)}
In this method, we maximize the non-Gaussianity given by
\begin{align}
J(Y) \propto \{E[G(Y)] - E[G(V)]\}^2,
\end{align}
where $Y$ is a random variable of zero mean and unit variance, which can be achieved by whitening the data, and $V$ is a Gaussian variable also of zero mean and unit variance. In function forms of $G_1$ and $G_2$ should not grow too fast to be robust. The following choice of $G$ is proposed by Hyvärinen and Oja.
\begin{align}
G_1(u) &= \frac{1}{a_1} \log (\cosh a_1 u). \\
G_1^{\prime}(u) &= \tanh(a_1 u).
\end{align}
We want to find the extrema of $E[G(Y)]$ to maximize $J(Y)$ under the constraints
\begin{align}
E[(w^T x)^2] = \| w\|^2 = 1.
\end{align}
From the Kuhn--Tucker conditions, the extremum condition is satisfied when
\begin{align}
E[x g(w^Tx)] - \beta w &= 0.
\end{align}
This equation can be solved by using Newton's method.
Let $F(w)$ denote the left-hand side of the equation, which we are trying to solve.
We update the value of $w$ in each iteration according to the following equation.
\begin{align}
w_{n+1} &= w_n - \frac{F(w_n)}{F^{\prime}(w_n)} \\
&= w_n - \frac{E[x g (w_n^T x)] - \beta w_n}{E[g^{\prime}(w_n^T x)] - \beta} \\
&= \frac{w_n E[g^{\prime}(w_n^T x)] - \beta w_n - E[x g (w_n^T x)] + \beta w_n}{E[g^{\prime}(w_n^T x)] - \beta} \\
&= \frac{w_n E[g^{\prime}(w_n^T x)] - E[x g (w_n^T x)]}{E[g^{\prime}(w_n^T x)] - \beta}.
\end{align}

\section{Applications}
1) sklearn example
3) example: sound single decomposition
4) stock price
5) EEG
% (https://towardsdatascience.com/independent-component-analysis-ica-a3eba0ccec35)

\section{How to check the robustness and }
% Amari metric, ref "ELements of Statistical Learning", p.570

>>>>

% You should use BibTeX and apsrev.bst for references
% Choosing a journal automatically selects the correct APS
% BibTeX style file (bst file), so only uncomment the line
% below if necessary.
%\bibliographystyle{apsrev4-2}
\usepackage{graphicx}
\usepackage{amsmath,amsthm,amssymb,mathtools}
\usepackage{amsfonts}
\usepackage{bm}
\usepackage{color}
\frenchspacing
%\graphicspath{{~/data/KannanLu/exp623/article/}}
\usepackage{listings}
\usepackage{color}
\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
% \definecolor{backcolour}{rgb}{0.18,0.18,0.18}
\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\footnotesize,
    breakatwhitespace=false,
    breaklines=true,
    captionpos=b,
    keepspaces=true,
    numbers=left,
    numbersep=5pt,
    showspaces=false,
    showstringspaces=false,
    showtabs=false,
    tabsize=2
}
\lstset{style=mystyle}

@@

\subsection{FastICA}
\subsection{alternative objective function, skewness and kurtosis (Kannan)}
In this method, we maximize the non-Gaussianity given by
\begin{align}
J(Y) \propto \{E[G(Y)] - E[G(V)]\}^2,
\end{align}
where $Y$ is a random variable of zero mean and unit variance, which can be achieved by whitening the data, and $V$ is a Gaussian variable also of zero mean and unit variance. In function forms of $G_1$ and $G_2$ should not grow too fast to be robust. The following choice of $G$ is proposed by Hyvärinen and Oja.
\begin{align}
G_1(u) &= \frac{1}{a_1} \log (\cosh a_1 u). \\
G_1^{\prime}(u) &= \tanh(a_1 u).
\end{align}
We want to find the extrema of $E[G(Y)]$ to maximize $J(Y)$ under the constraints
\begin{align}
E[(w^T x)^2] = \| w\|^2 = 1.
\end{align}
From the Kuhn--Tucker conditions, the extremum condition is satisfied when
\begin{align}
E[x g(w^Tx)] - \beta w &= 0.
\end{align}
This equation can be solved by using Newton's method.
Let $F(w)$ denote the left-hand side of the equation, which we are trying to solve.
We update the value of $w$ in each iteration according to the following equation.
\begin{align}
w_{n+1} &= w_n - \frac{F(w_n)}{F^{\prime}(w_n)} \\
&= w_n - \frac{E[x g (w_n^T x)] - \beta w_n}{E[g^{\prime}(w_n^T x)] - \beta} \\
&= \frac{w_n E[g^{\prime}(w_n^T x)] - \beta w_n - E[x g (w_n^T x)] + \beta w_n}{E[g^{\prime}(w_n^T x)] - \beta} \\
&= \frac{w_n E[g^{\prime}(w_n^T x)] - E[x g (w_n^T x)]}{E[g^{\prime}(w_n^T x)] - \beta}.
\end{align}
\section{Example}
We implement FastICA according to the formalism described in the previous section.
The algorithm is applied to a following set of signal functions.
\begin{align}
s_1 &= \sin(2 t). \\
s_2 &= \mathrm{sign} [\sin(3 t)]. \\
s_3 &= \frac{1}{2\pi} (t~\mathrm{mod}~2\pi).
\end{align}
The mixing matrix is taken to be
\begin{align}
A &= \begin{bmatrix}
1 & 1 & 1 \\
0.5 & 2 & 1 \\
1.5 & 1 & 2
\end{bmatrix}.
\end{align}
The result from our implementation and FastICA are shown in Fig.~\ref{fig:sklearn}.
\begin{figure}
\includegraphics{scripts/sklearn.pdf}
\caption{\label{fig:sklearn} FastICA results using our code and \lstinline{sklearn.decomposition.FastICA}}
\end{figure}

\section{Applications}
3) example: sound single decomposition
4) EEG % (https://towardsdatascience.com/independent-component-analysis-ica-a3eba0ccec35)
5) stock price

\section{How to check the robustness and convergence}
% Amari metric, ref "ELements of Statistical Learning", p.570
